Bandit Multiclass Linear Classification: Efficient Algorithms for the Separable Case

### 2019-01-22 ICML-submission/empirical-comparison.tex
Fix experiments section

\section{Experiments}
\label{sec:experiments}

We compare our algorithm with \textsc{Banditron} empirically to show that our
kernelized bandit algorithm can have significant advantage over
\textsc{Banditron} in the separable case.

We generated strongly and weakly separable samples with $K=3$ classes shown
in~\autoref{figure:strongly-and-weakly-separable-datasets}. The data sets are
shown as two-dimensional, however, we used a third dimension (so-called
\emph{constant} feature) i.e. a feature that for every example is set to $1$. We
added the constant feature so that \autoref{figure:strongly-separable-dataset}
is actually strongly lineraly separable. For both data sets, the sample size is
$T=5\times 10^6$ and the margin is $\gamma=0.05$.

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.23\textwidth}
\captionsetup{justification=centering}
\begin{center}
\hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_points}
\caption{Strongly separable case}
\label{figure:strongly-separable-dataset}
\end{center}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.23\textwidth}
\captionsetup{justification=centering}
\centering
\hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_points}
\caption{Weakly separable case}
\label{figure:weakly-separable-dataset}
\end{subfigure}
\vspace*{-0.2cm}
\caption{Strongly and weakly generated data sets with $K=3$ classes. All examples lie in
the unit ball. Class 1 is depicted red. Classes 2 and 3 are depcited gree and
blue, respectively. $80\%$ of example belong to class 1, $10\%$ of the data
belong to class 2 and $10\%$ belong to class 3. Class 1 occupies the angle
interval $[-15^\circ, 15^\circ]$, while classes 2 and 3 occupy angle intervals
$[15^\circ, 180^\circ]$ and $[-180^\circ, -15^\circ]$ respectively. We removed
points that lie within margin $\gamma=0.05$ of the linear separators.}
\label{figure:strongly-and-weakly-separable-datasets}
\end{figure}

All results are averaged over $20$ runs.
We plot the number of mistakes against time in Figure~\ref{fig:both}. We can see
there is indeed a dilemma for \textsc{Banditron}: when picking a relatively
large exploration rate, it suffers from the cost of continuous exploration; when
picking a relatively small exploration rate, it cannot adapt its model quickly
enough (this problem exacerbates when a majority class occupies a small region,
like class 1). Striking a balance between the two extremes leads to the
$\sqrt{T}$ mistake bound. On the contrary, our algorithm makes significant
progress on every mistake it makes, leading to a finite mistake bound.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \begin{center}
        \hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0.1cm, 0, 0}, clip]{figures/strong3}
        \caption{Strongly separable case}
        \end{center}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.23\textwidth}
        \captionsetup{justification=centering}
        \centering
        \hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0.1cm, 0, 0}, clip]{figures/weak3}
        \caption{Weakly separable case}
    \end{subfigure}
    \vspace*{-0.2cm}
    \caption{Comparison between our algorithm and Banditron with different exploration parameter $\epsilon$ under $\gamma=0.05$ and $K=3$.}
    \label{fig:both}
\end{figure}

>>>>

\section{Experiments}
\label{section:experiments}

We implemented
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
Algorithm~\ref{algorithm:kernelized} with
kernel~\eqref{equation:rational-kernel}, and the \textsc{Banditron} algorithm.
We generated strongly and weakly separable data sets with $K=3$ classes. The
data sets are shown in~\autoref{figure:strongly-and-weakly-separable-datasets}.
Although, the data sets are shown as two-dimensional, we used a third dimension,
the so-called \emph{constant feature}, i.e. a feature that for every example is
set to $1$. We added the constant feature so that
\autoref{figure:strongly-separable-dataset} is actually strongly lineraly
separable. For both data sets, the sample size is $T=5\times 10^6$ and the
margin is $\gamma=0.05$.

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.23\textwidth}
\captionsetup{justification=centering}
\begin{center}
\hspace*{-0.3cm} \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/strong_points}
\caption{Strongly separable case}
\label{figure:strongly-separable-dataset}
\end{center}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.23\textwidth}
\captionsetup{justification=centering}
\centering
\hspace*{-0.3cm}  \includegraphics[width=1.15\textwidth, trim={0, 0cm, 0, 0}, clip]{figures/weak_points}
\caption{Weakly separable case}
\label{figure:weakly-separable-dataset}
\end{subfigure}
\vspace*{-0.2cm}
\caption{Strongly and weakly generated data sets with $K=3$ classes. All
examples lie in the unit ball of $\R^2$. Class 1 is depicted red. Classes 2 and
3 are depcited gree and blue, respectively. $80\%$ of the examples belong to class 1,
$10\%$ belong to class 2 and $10\%$ belong to class 3. Class 1
occupies the angle interval $[-15^\circ, 15^\circ]$, while classes 2 and 3
occupy angle intervals $[15^\circ, 180^\circ]$ and $[-180^\circ, -15^\circ]$
respectively. We removed points that lie within margin $\gamma=0.05$ of the
linear separator.}
\label{figure:strongly-and-weakly-separable-datasets}
\end{figure}

We evaluated all three the algorithms on the two data sets. Since all three
algorithms are randomized, we run each algorithm $20$ times. \textsc{Banditron}
has exploration rate parameter $\epsilon$ for which we tried values $\epsilon
\in \{0.02, 0.01, 0.005, 0.002, 0.001, 0.0005 \}$. The average cumulative number
of mistakes up to round $t$ as a function $t$ are shown in
Figures~\ref{figure:number-of-mistakes-strongly-separable-dataset}
and~\ref{figure:number-of-mistakes-weakly-separable-dataset}.

We can see there is a dilemma for \textsc{Banditron}. When the
exploration rate $\epsilon$ is large, it suffers from the cost of exploration.
When the exploration rate is small, it cannot adapt its model quickly enough.
This problem is exacerbated when a majority class occupies a small region, like
class 1. Striking a balance between the two extremes leads to the
$\Theta(\sqrt{T})$ number of mistakes. In contrast, our algorithm makes
significant progress on every mistake it makes, leading to a finite mistake
bound.

Appendix~\ref{section:supp-to-experiment} shows decision boundaries
that each of the algorithms learns.

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/strong3}
\caption{Strongly separable case. Cumulative number of mistakes of various algorithms.}
\label{figure:number-of-mistakes-strongly-separable-dataset}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/weak3}
\caption{Weakly separable case. Cumulative number of mistakes of various algorithms.}
\label{figure:number-of-mistakes-weakly-separable-dataset}
\end{figure}

### 2019-01-22 ICML-submission/from-weak-separability-to-strong-separability.tex
Fix experiments section

The kernel function $k:\B(\zero,1) \times \B(\zero,1) \to \R$ is defined as
$$
k(x,x') = \frac{1}{1 - \frac{1}{2}\ip{x}{x'}_{\R^d}} \; .
$$
Note that kernel function $k(x,x')$ can be evaluated in $O(d)$ time.

@@

Using the lemma and the polynomial approximation theorems, we can prove that the
mapping $\phi$ maps any set of weakly linearly separable examples to a strongly
linearly separable set of examples. Due to space constraints, we only sketch the
proof of Theorem~\ref{theorem:margin-transformation} here, and defer the full
proof to Appendix~\ref{section:full-proof-of-theorem-margin-transformation}.

>>>>

The kernel function $k:\B(\zero,1) \times \B(\zero,1) \to \R$ is defined as
\begin{equation}
\label{equation:rational-kernel}
k(x,x') = \frac{1}{1 - \frac{1}{2}\ip{x}{x'}_{\R^d}} \; .
\end{equation}
Note that kernel function $k(x,x')$ can be evaluated in $O(d)$ time.

@@

Using the lemma and the polynomial approximation theorems, we can prove that the
mapping $\phi$ maps any set of weakly linearly separable examples to a strongly
linearly separable set of examples. Due to space constraints, we defer the full
proof to Appendix~\ref{section:proof-of-theorem-margin-transformation}.
